---
title: "Gender and womens' protests"
author: "James Goldie, 360info"
---

```{r setup}
library(tidyverse)
library(themes360info)
library(ggforce)
library(nVennR)
library(here)
```

```{r load}

news <- read_tsv(here("data", "worldprotests", "curator8.tsv"))

# tidy the themes coding file
read_csv(here("data", "worldprotests", "machine-readable.csv")) %>%
  select(
    id_num = INC_ID_,
    id_country = CNT_NAM,
    id_end_month = DAT_ENM,
    id_end_year = DAT_ENY,
    id_start_month = DAT_STM,
    id_start_year = DAT_STY,
    count_arrested =  REP_ARN,
    count_dead = REP_DEN,
    count_displaced = REP_DSN,
    count_injured = REP_IJN,
    count_missing = REP_MSN,
    everything()) %>%
  # give remaining columns a prefix
  rename_with(~ paste0("theme_", .x), !starts_with(c("id_", "count_"))) %>%
  # convert theme cols to logical
  mutate(across(starts_with("theme_"), as.logical)) ->
themes

# map of column names to long names
theme_dict <-
  read_csv(here("data", "worldprotests", "categories.csv")) %>%
  select(code, en)
```

Okay. Let's look at events that are coded as being related to themes "Women's/Girls' Rights" `RGT_WMN` or "Women/Girls/Feminist" `GRP_WGF`. What other themes emerge from them?

```{r femevents}

# we're going to filter on incidents where women and girls' rights themes
# are observed, then lengthen out by theme to see what other themes are
# present
themes %>%
  select(-ends_with("month"), -starts_with("count_")) %>%
  filter(theme_RGT_WMN | theme_GRP_WGF) ->
womens_themes_wide

womens_themes_wide %>%
  pivot_longer(starts_with("theme_"),
    names_to = "theme", values_to = "present") %>%
  filter(present) %>%
  # now pull in the long theme names and order them by prevalence
  mutate(
    theme = str_replace_all(theme, "theme_", ""),
    # now turn the theme into a factor to get the long labels
    theme_long = factor(theme, levels = theme_dict$code, labels = theme_dict$en),
    theme_long = fct_rev(fct_infreq(theme_long))) %>%
  select(-present) ->
womens_themes_long
  
# now calculate counts and proportions
womens_themes_long %>%
  group_by(theme, theme_long) %>%
  summarise(
    count = length(theme_long),
    prop = count / nrow(womens_themes_wide)) %>%
  ungroup() %>%
  filter(!is.na(theme_long))  ->
womens_themes_freq

theme_freq_barplot <-
  ggplot(womens_themes_freq) +
  aes(x = theme_long, y = prop) +
  geom_col(fill = "black") +
  coord_flip() +
  scale_y_continuous(labels = scales::label_percent(), position = "right") +
  labs(
    x = NULL, y = "Proportion of articles",
    title = toupper("Themes in protests"),
    subtitle = toupper("Subtitle here"),
    caption = "**Source:** worldprotests.org") +
    theme_360(base_size = 14)

save_360plot(theme_freq_barplot, here("out", "theme_freq_barplot.png"), shape = "phone-portrait")

```

Let's focus in on:

MET_*: Type of Action
GRV_*: Type of grievance
GJU_*: ?
OPP_*: Type of opponents
EPL_*: ?
REP_*: Repercussions (some are counts, or NA if unavilable)
RGT_*: Rights demanded?
SCP_*: Spatial scope
REG_*: Region
ICM_*: Income bracket (of whom? probably country income)

Actors:
  * GRP_WGF: Women/Girls/Feminist
  * GRP_IND: Indigenous Group
  * GRP_ERG: Ethnic/Racial Group

Rights demanded:
  * RGT_WMN: Women's/Girls' Rights
  * RGT_GAY: LGBT/Sexual Rights 
  * RGT_EIR: Ethnic/Indigenous/Racial Justice
  * RGT_ENV: Environment and the Commons
  * RGT_IMM: Immigrant Rights
  * RGT_DEN: Deny Rights/Reject Equal Rights for a Group

* Other aspects
  * ACH_ENV: Environmental ?
  * GJU_ENV: Environment/Climate Justice
  * GJU_IMP: Anti-Imperialism


Let's go back to the original dataset (all he events, not just the subsetted ones) and pick out the ones that include _any_ of the following themes.

```{r test}
themes %>%
  select(-starts_with("count_")) %>%
  # filter for only the events that have one of these themes
  filter(if_any(
    all_of(c("theme_GRP_WGF", "theme_GRP_IND", "theme_GRP_ERG",
      "theme_RGT_WMN", "theme_RGT_GAY", "theme_RGT_EIR", "theme_RGT_ENV",
      "theme_RGT_IMM", "theme_RGT_DEN")),
    ~ .x)) ->
event_subset
```

### Group associations

Which groups turn up to certain kinds of events? Either (a) events that other groups have turned up to, or (b) events about something in particular?

```{r assoc}

# find_associations: tell me how often other themes featured in protests
# with a basis theme?
# - df: data farme of events
# - basis: bare name of the column to subset to first
# - category: string of category of themes to look for associations in (eg. "GRP")

find_associations <- function(df, basis, category) {

  df %>%
    filter({{ basis }}) ->
  df_filtered

  # get max count first!
  count_all <- nrow(df_filtered)
    
  df_filtered %>%
    # calculate counts for the groups
    summarise(across(
      starts_with(paste0("theme_", category, "_")),
      ~ length(which(.x)))) %>%
    # lengthen and get theme descriptions
    pivot_longer(
      starts_with(paste0("theme_", category, "_")),
      names_to = c("association"),
      values_to = "count") %>%
    mutate(association = str_replace(association, "theme_", "")) %>%
    left_join(theme_dict, by = c("association" = "code")) %>%
    filter(count > 0) %>%
    # proportion calculated
    mutate(prop = count / count_all)
}

event_subset %>% find_associations(theme_GRP_WGF, "GRP") -> womens_assoc
event_subset %>% find_associations(theme_GRP_IND, "GRP") -> indig_assoc
event_subset %>% find_associations(theme_GRP_ERG, "GRP") -> racial_assoc
event_subset %>% find_associations(theme_RGT_GAY, "GRP") -> queer_assoc
event_subset %>% find_associations(theme_GJU_ENV, "GRP") -> clim_assoc

all_assoc <- bind_rows(
  list(
    "... with\nwomens'\ngroups?" = womens_assoc,
    "... with\nindigenous\ngroups?" = indig_assoc,
    "... with\nethnic and racial\ngroups?" = racial_assoc,
    "... for\nqueer\nrights?" = queer_assoc,
    "... for\nenvironment\nor climate\njustice?" = clim_assoc
    ),
  .id = "basis")

# plot the associations
all_assoc %>%
  mutate(
    en = str_replace(en, " ", "\n"),
    # highlight = if_else(
    #   group %in% c("GRP_IND", "GRP_ERG", "GRP_WGF"),
    #   pal_360[["blue"]],
    #   pal_360[["black"]])
  ) %>%
  {
    ggplot(.) +
      aes(x = en, y = count) +
      geom_col() +
      scale_fill_identity() +
      coord_flip() +
      facet_row(vars(basis)) +
      theme_360(base_size = 12) +
      theme(
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()
      ) +
      labs(
        x = NULL, y = "Number of events",
        title = toupper("Group associations"),
        subtitle = toupper("Which groups appear at protests..."))
  } %>%
  print() %>%
  save_360plot(here("out", "associations.png"), shape = "square")

```

### Actions

Which kinds of protest method

```{r opponents}
event_subset %>% find_associations(theme_GRP_WGF, "MET") -> womens_met
event_subset %>% find_associations(theme_GRP_ERG, "MET") -> racial_met
event_subset %>% find_associations(theme_GJU_ENV, "MET") -> clim_met

all_met <-
  bind_rows(
    list(
      "... with **womens' groups?**" = womens_met,
      # "... with indigenous groups?" = indig_met,
      "... with **ethnic and racial groups?**" = racial_met,
      # "... for queer rights?" = queer_met,
      "... for **environment or climate justice?**" = clim_met
      ),
    .id = "basis") %>%
  # add missing protest methods
  complete(basis, en, fill = list(count = 0, prop = 0)) %>%
  # remove violent methods since we're focusing on non-violent protest
  filter(!(association %in% c("MET_VIO", "MET_VAL"))) %>%
  # order b=ars by total frequency
  mutate(en = fct_reorder(str_replace(en, " ", "\n"), count))

# plot the opponents
all_met %>%
  {
    ggplot(.) +
      aes(x = en, y = count, fill = basis) +
      geom_col(position = "dodge", width = 0.6) +
      scale_y_continuous(expand = expansion(0, 0)) +
      scale_fill_manual(
        values = c(
          pal_360[["green"]],
          pal_360[["darkblue"]],
          pal_360[["blue"]]),
        guide = guide_legend(reverse = TRUE)) +
      coord_flip() +
      # facet_row(vars(basis)) +
      theme_360(base_size = 16) +
      theme(
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend position and sizing
        legend.position = "top",
        legend.direction = "vertical",
        legend.justification = "left",
        legend.margin = margin(6, 6, 6, 0),
        legend.key.width = unit(0.6, "cm"),
        legend.key.height = unit(0.6, "cm"),
        legend.text = element_textbox(
          size = rel(1.1),
          width = unit(13, "cm"))
      ) +
      labs(
        x = NULL, y = "Number of events", fill = NULL,
        title = toupper("Non-violent Protest methods"),
        subtitle = toupper("Which non-violent methods are used at protests...")) +
      labs(caption = paste(
        "**CHART:** James Goldie, 360info",
        "**DATA:** worldprotests.org",
        sep = "<br>"))
  } %>%
  print() %>%
  save_360plot(here("out", "methods.png"), shape = "phone-portrait")
```